---
title: "XAJ_modelling"
author: "Xiaoxiang Guan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This is a brief tutorial on how to implement the source code for hydrological modelling,
by using Xinanjiang (XAJ) model. 

First, load the `XAJ.R` source code into the console (R session), together with the `rtop` 
library in order to use the SCE-UA algorithm for model calibration. 
```{r load, echo=TRUE, warning=FALSE, message=FALSE}

library(rtop)
ws <- "D:/HydroModel/"
source(paste0(ws, "scr/XAJ.R"))

```

## Import the example data

import the example data `data/DATA_Area4766_1982-2002.csv`. This example data set covers 
the basic variables needed for modelling: precipitation (mm), potential evapotranspiration (mm) and 
discharge (m3/s) all at the daily step. The dataset covers the observation period of 1982-2002.
The example catchment has an area of 4766 km2.

```{r example_data, echo=TRUE}
dat = read.table(paste0(ws, "data/DATA_Area4766_1982-2002.csv"),header = T,sep = ",")

P = dat[,4]
E = dat[,5]
Q = dat[,6]
DAREA = 4766 ##
DT = 24
head(dat, 8)

```

## XAJ model calibration based on SCE-UA algorithm

create the global variables in order to make the data available for the 
calibration (fitting) function during mutiple iteration processes. 

```{r Global_variables, echo=TRUE}
#-----定义全局变量 (define global variables)-----
Validinput_P <<- P
Validinput_E <<- E
Validinput_Q <<- Q
Validinput_DAREA <<- DAREA  ## catchment area in km2
Validinput_DT <<- 24  ## simulation time step

```

there are 15 parameters in XAJ we defined to be calibrated. 
For the description on the parameters, refer to the documents.

```{r XAJ_para_boundary, echo=TRUE}
#-------------1    2    3     4    5     6     7    8     9    10   11   12   13   14  15
#-------------Kc, SM,   B,    C   Ex   WUM   WLM  WDM    Kg    Ki   Cs   Ci   Cg   CR  L
paralower = c(0.1,5  ,  0.1, 0.05, 1.0,  40,  20,  5, 0.25, 0.25, 0.1, 0.8, 0.8, 0.1, 0)
paraupper = c(1  ,50 ,  1.5,  0.2, 1.4, 300, 150, 60, 0.49, 0.49,   6,   1,   1, 0.6, 3)
parainitial=c(0.6,20 ,  0.4,  0.1, 1.2, 180,  40, 10,  0.3,  0.4, 0.4, 0.9, 0.9, 0.3, 0)
fit = function(x) {
  Kc = x[1]; SM = x[2];WUM = x[6];WLM = x[7];WDM = x[9];
  B = x[3];C = x[4];Ex = x[5];Kg = x[9];Ki = x[10];
  Cs = x[11];Ci = x[12];Cg = x[13];CR = x[14];L = x[15]
  out = XAJ(Validinput_P, Validinput_E, Validinput_Q, Validinput_DAREA, Validinput_DT,
            Kc, SM, B, C,Ex,WUM, WLM, WDM, 
            Kg, Ki, 
            Cg, Cs, Ci, CR,L) 
  Qsi = out
  Qre = Validinput_Q
  R2 = 1-sum((Qre-Qsi)^2,na.rm = T)/sum((Qre-mean(Qre))^2,na.rm = T) #NSE
  return(1-R2)
}

```

calibrate the XAJ model by using the `sceua()` from R package `rtop`. The calibrated parameters
(results) could be used to rerun the model to check the performance of model.

```{r model_calibration, echo=TRUE, error=FALSE,warning=FALSE,message=FALSE}
sceua_xaj = sceua(fit, pars = parainitial, lower = paralower, 
                  upper = paraupper, maxn = 1000, pcento = 0.00001, iprint = 0)
x = sceua_xaj$par
print(x)
```


```{r, comparison}

Kc = x[1]; SM = x[2];WUM = x[6];WLM = x[7];WDM = x[9];
B = x[3];C = x[4];Ex = x[5];Kg = x[9];Ki = x[10];
Cs = x[11];Ci = x[12];Cg = x[13];CR = x[14];L = x[15]
out = XAJ(Validinput_P, Validinput_E, Validinput_Q, Validinput_DAREA, Validinput_DT,
          Kc, SM, B, C,Ex,WUM, WLM, WDM, 
          Kg, Ki, 
          Cg, Cs, Ci, CR,L) 
Qsi = out
Qre = Validinput_Q
R2 = 1-sum((Qre-Qsi)^2,na.rm = T)/sum((Qre-mean(Qre))^2,na.rm = T) #NSE
print(paste0("NSE: ", round(R2, 4)))

#--- obs VS. sim
plot(1:length(Validinput_Q),Validinput_Q, 
     xlab = "time", ylab = "discharge [m3/s]", main = "simulation vs. observation")
points(1:length(Qsi),Qsi, type = "l",col = "red") 

```
here we plot the observed and simulated discharge processes in the same plot 
(observation: black points, and simulation: red line), with an 
attempt to have a straightforward understanding how well the XAJ simulates the discharge 
processes. The result displays that the XAJ model performs well in discharge simulation 
for the example catchment, under the support the SCE-UA algorithm.
The main conclusion could be the XAJ model together with the parameter calibration algorithm 
shows a good performance and have a positive application potential. 

